### 2020-09-03 Manuscript/What.lies.ahead.reviewed.tex
edited and updated up to line number 170 in the tex file, which is line 226 in the PDF

By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time \cite{kelly2019understanding}, using pooled data from technical replicates and mean proportions across biological replicates. {\color{red} RGS: We used this index of abundance to calculate the multivariate community analysis, while species responses used the presence/absence dataset.  RPK: which analyses use the index, rather than presence-absence? make that clear here.}

@@

To model biological responses to present and future scenarios, we used a hierarchical logistical regression model relating the presence of each taxon to temperature and pH, in which the slopes of temperature and pH effects varied by taxon, and each taxon had a unique intercept that was allowed to vary by geographic area. For each taxon, we fit these models using the Bayesian generalized linear mixed effects functions in R package rstanarm \cite{rstanarm} for R. Model selection using WAIC supported this as the preferred model over several similar ones (see Supporting Information).

Given the estimated sea-surface temperatures and pH values for 2095 and observed data from 2017, we then used taxon-specific logistic regression models to estimate the suitability of habitat for each taxon in the future scenario. For each point in the pH - temperature grid, we calculated species richness as the mean result of 100 independent draws of the posterior distributions of all modeled taxa.

@@

%%%%% Figure 1: map and environmental conditions
\begin{figure}%[tbhp]
\centering
\includegraphics[width=1 \linewidth]{Figures_for_MS/Fig1.png}
\caption{\footnotesize {\textbf{A} Sampling locations in intertidal areas of the Hood Canal ({\color{red}dark gold}) and San Juan Island ({\color{red}turquoise}). \textbf{B} Planktonic richness (unique taxa) per sampling locality, as reflected by the eDNA COI assay.{\color{red}\textbf{B.1} Taxa from all phyla; \textbf{B.2} Diatoms(above) and dinoflagellates (below) shown separately} \textbf{C} The observed environmental profiles of these two regions reflect a broad range of environmental conditions, with the Hood Canal resembling future conditions in temperate areas worldwide. Shown: Dissolved Inorganic Carbon (DIC, $\mu$moles/kg); Total Alkalinity (TA, $\mu$moles/kg); Temperature ($\degree$C) and Salinity (PSU). Individual sampling sites shown as facets. {\color{red} CHANGE y-scale for Dinophyceae DONE}}}
\label{fig:Fig1}
\end{figure}
%%%%%

@@

Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to geographic Area (Hood Canal vs. San Juan Island; F-model = 5.4; p $<$ 0.0001). {\color{red} RPK: still up-to-date? no adonis code in the Rmd file that I saw. RGS: that's true - I added new chunk of code line 382 \& updated results}. We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig. 2 shows results for Hood Canal; full analysis in the Supporting Information). 
{\color{red}\st{Independent of environmental parameters, we separately clustered samples by pairwise Bray-Curtis dissimilarities among metabarcoding communities (k-means; N = 3) to identify groups of samples that were similar to one another with respect to biological community.}}
 Each biological cluster (colored hulls, Fig. 2.A) occupied a unique area of environmental parameter space. Planktonic communities therefore varied predictably with water temperature, salinity, and pH, across a range of those parameters likely to be encountered in many near-term future-ocean scenarios \cite{murray2015inland}. 
 Multinomial logistic regression yielded predictions of the most-likely community for any combination of environmental parameters (Fig. 2B).

{\color{red} RPK: I think everything below needs updating, according to some final analysis. TODO: RGS}

These communities were distinguished by changes in the relative abundances of many taxa: in the Hood Canal for example (Fig. 2C), the colder water and higher pH communities showed higher values of \textit{Centropages} sp., \textit{Nitzschia longissima}, \textit{Prorocentrum} sp., \textit{Chrysochromulina} sp., \textit{Navicula} sp. and \textit{Chaetoceros} sp. Warmer waters were dominated by \textit{Bathycoccus prasinos} and \textit{Cymbomonas} sp., while at lower pH values, marine communities showed \textit{Litchtheimia ramosa}, \textit{Micromonas pusilla} at higher relative abundances.

\underline{Climate envelopes and future distributions}

The low-pH and {\color{red}high-temperature} planktonic communities (cluster 3, Fig. 2{\color{red}A and B}) occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls. For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12 and pH $<$ 8.2 -- where cluster 3 is the most likely community -- to occur more than two-thirds of the time (68\%).

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{	Figures_for_MS/Figure2.2020-08-27.png}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model {\color{red}(salinity was not represented as it was tightly correlated with temperature)}. \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supporting Information for full analysis.}}
\label{fig:Fig2}
\end{figure}
%Move this to the methods%
 To explore the suitability of different environmental conditions for each taxon, we modeled the likelihood of taxon presence as a function of area and season (random effects) and temperature and pH (fixed effects).  Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}. Using predicted conditions for Puget Sound for the year 2095 \cite{khangaonkar2019salish}, we then projected the probability of presence of each taxon across these conditions over 100 draws of the posterior probabilities of all taxa. We used these draws to 1) show the change in the probability of presence of each taxon for each scenario (i.e. 2017 vs 2095, Fig.3 top panel); 2) calculate the  taxon richness within taxonomic groups across the pH-temperature continuum (Fig. 3 center panel),% 2) show the consequent variation in richness for higher taxa across the pH - temperature continuum (Figure 3B),
  and 3) visualized the change in suitability for each taxon in our dataset individually (selected taxa in Fig. 3 bottom panel).

>>>>

By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time \cite{kelly2019understanding}, using pooled data from technical replicates and mean proportions across biological replicates. We used this index of abundance in the multivariate community analysis, and used binary (presence/absence) data to capture species-level responses to environmental conditions.

@@

To model biological responses to present and future scenarios, we used a hierarchical logistical regression model relating the presence of each taxon to temperature and pH, in which the slopes of temperature and pH effects varied by taxon, and each taxon had a unique intercept that was allowed to vary by geographic area. For each taxon, we fit these models using the Bayesian generalized linear mixed effects functions in R package rstanarm \cite{rstanarm} for R. Model selection using WAIC \cite{watanabe2010asymptotic} supported this as the preferred model over several similar ones (see Supporting Information for model comparison information and code) and helped to avoid model overfitting and maintain out-of-sample predictive power.

Given the sea-surface temperatures and pH values for 2017 (observed) and 2095 (estimated) and taxon-specific logistic regression models, we then evaluated the suitability of habitat for each taxon in the future scenario. For each point in the pH - temperature grid, we calculated a species' probability of presence as the mean of 100 independent draws of the posterior model response. We estimated species richness as the sum of taxa with $>$ 50\% probability of occurrence at a given set of environmental parameters.

@@

%%%%% Figure 1: map and environmental conditions
\begin{figure}%[tbhp]
\centering
\includegraphics[width=1 \linewidth]{Figures_for_MS/Fig1.png}
\caption{\footnotesize {\textbf{A} Sampling locations in intertidal areas of the Hood Canal (dark gold) and San Juan Island (turquoise). \textbf{B} Planktonic richness (unique taxa) per sampling locality, as reflected by the eDNA COI assay. Boxplots represent the variability in richness across all time points for samples taken at the indicated site. \textbf{B.1} Taxa from all phyla; \textbf{B.2} Diatoms(above) and dinoflagellates (below) shown separately; note change in the scale of the y-axis. \textbf{C} The observed environmental profiles of these two regions reflect a broad range of environmental conditions, with the Hood Canal resembling future conditions in temperate areas worldwide. Shown: Dissolved Inorganic Carbon (DIC, $\mu$moles/kg); Total Alkalinity (TA, $\mu$moles/kg); Temperature ($\degree$C) and Salinity (PSU). Individual sampling sites shown as vertical facets.}}
\label{fig:Fig1}
\end{figure}
%%%%%

@@

Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to geographic Area (Hood Canal vs. San Juan Island; F-model = 5.4; p $<$ 0.0001). We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig. 2; for clarity, results for Hood Canal shown; full analysis in the Supporting Information).

 Each biological cluster (colored hulls, Fig. 2.A) occupied a unique area of environmental parameter space. Planktonic communities therefore varied predictably with water temperature, salinity, and pH, across a range of those parameters likely to be encountered in many near-term future-ocean scenarios \cite{murray2015inland}. 
 Multinomial logistic regression yielded predictions of the most-likely community for any combination of environmental parameters (Fig. 2B).

These communities were distinguished by changes in the relative abundances of a wide variety of taxa. In the Hood Canal for example, the colder water and higher-pH community (community 1 in Fig. 2C) showed higher eDNA indexes of diatoms \textit{Thalassiosira nordenskioeldii}, \textit{Nitzschia cf. promare}, and \textit{Ditylum brightwellii} -- among other species -- relative to a higher-temperature community (community 3 in Fig. 2C). Communities 2 and 3 had relatively higher values for taxa from a mix of groups including \textit{Emiliana huxleyi}, \textit{Minutocellus polymorphus}, \textit{Micromonas pusilla}, and \textit{Bathycoccus prasinos}. Community 3 (Fig. 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls. For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12 and pH $<$ 8.2 -- where community 3 is the most likely community -- to occur more than two-thirds of the time (68\%). {\color{red} RAMON -- can you see if this 68\% stat is still true, and also if you can incorporate the similar expectations that are commented-out directly below, updating their \% if necessary?}  \\

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{	Figures_for_MS/Figure2.2020-08-27.png}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model (salinity was uninformative, being tightly correlated with temperature). \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supporting Information for full analysis.}}
\label{fig:Fig2}
\end{figure}

\noindent\underline{Climate envelopes and future distributions}

To explore the suitability of different environmental conditions for each taxon, we modeled the likelihood of taxon presence as a function of temperature, pH, and geographic Area as described in the Methods. Salinity was not informative in our models, as it is highly correlated with temperature in our dataset and is moreover predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}. Model projections let us show the change in the probability of presence of each individual taxon for 2095 vs 2017 (Fig. 3 A.1, B.1, C.1), estimate richness for larger taxonomic grounds as a whole for these two climates (Fig. 3 A.2, B.2, C.2), and estimate richness within taxonomic groups across the pH-temperature continuum (Fig. 3 A.3, B.3, C.3).

{\color{red} RAMON - I edited everything above this point. From here down, the text all needs updating to reflect the results in the current figures. }
